---
title: "Monitoring Assessment - Literature Corpus"
author: Jessica Hetzer <jessica.hetzer@senckenberg.de> 
format: html
code-fold: true
editor: visual
editor_options: 
  chunk_output_type: inline
---

```{r,  echo=FALSE, warning=FALSE, message=FALSE}

#load packages
Sys.setenv(openalexR.message = "suppressed")
library(openalexR)
library(jsonlite)
library(DT)
library(dplyr)


```

# Expanding core knowledge of authors

In this report, we introduce a method to create a comprehensive literature corpus by tracing networks of cited knowledge. We start with **key references provided by the authors** and link them to additional metadata (using openalexR). From these initial references, the search is expanded to include related publications through a network-based approach (snowballing, using the oa_snowball() function). This results in a broader collection of highly relevant literature.

Unlike standard keyword searches in databases like Web of Science, this approach builds a curated, up-to-date resource of key publications **directly related to the authors' field** and their chapters. The full corpus is saved and **can be filtered and searched,** as well as **explored interactively**, offering an intuitive way to browse and discover relevant research.

## Chapter 1

```{r}

fn <- file.path(".", "data", "1_Chapter","chapter1_seed_works.rds")
if (file.exists(fn)) {
    key_works <- readRDS(fn)
} else {
    seeds <- fromJSON(file.path(".", "input", "1_Chapter","chapter1_seed.json"))
    #seeds$DOI[is.na(seeds$DOI)]  <- seeds$URL[is.na(seeds$DOI)] 
    key_works <- oa_fetch(
        entity = "works",
        doi=seeds$DOI[!is.na(seeds$DOI)],
        verbose = TRUE
    )
    saveRDS(key_works, fn)
}

ids <- openalexR:::shorten_oaid(key_works$id)

fn <- file.path(".", "data", "1_Chapter", "chapter1_snowball.rds")
if (file.exists(fn)) {
    snowball <- readRDS(fn)
} else {
    snowball <- oa_snowball(
        identifier = ids,
        verbose = TRUE
    )
    saveRDS(snowball, fn)
}

```

Number of key references: 154

Size of corpus: \~ 23,000

*Please note that the following table is not complete yet. It is an intermediate product for demonstration purposes.*

```{r, echo=FALSE, warning=FALSE}

  snowball$nodes$author_short <- sapply(snowball$nodes$author, function(authors) {
    if (is.data.frame(authors)) {
      paste(authors$au_display_name[1], "et al.")
    } else if (is.character(authors)) {
      authors
    } else {
      NA_character_
    }
  })

# Create key_data_df from the snowball object
key_data_df <- data.frame(
  Authors = snowball$nodes$author_short,
  Year = snowball$nodes$publication_year,
  Title = snowball$nodes$title,
  Source = paste0(
    '<a href="',
    snowball$nodes$url,
    '" target="_blank">',
    snowball$nodes$url,
    '</a>'
  ),
  #Is_Key = snowball$nodes$is_key,
  stringsAsFactors = FALSE
)

# Display the filtered table using tableDT
# Display the filtered data using DT
datatable(
  key_data_df,
  escape = FALSE,
  options = list(
    pageLength = 10,
    searchHighlight = TRUE,
    dom = 'ftip'
  )
)
```

```{r, echo=FALSE, eval=FALSE}

library(RColorBrewer)
library(networkD3)
library(dplyr)
plot_snowball_interactive <- function(snowball, key_works, file = "interactive_graph_with_legend.html") {
  networkData <- data.frame(
    src = snowball$edges$from,
    target = snowball$edges$to,
    stringsAsFactors = FALSE
  )
  
  nodes <- data.frame(
    name = snowball$nodes$id,
    author = snowball$nodes$author_short,
    doi = snowball$nodes$doi,
    nodesize = snowball$nodes$cited_by_count / (2024 - snowball$nodes$publication_year) * 0.5,
    group = sapply(snowball$nodes$topics, function(x) {
      if ("display_name" %in% colnames(x)) {
        return(x$display_name[4])
      } else {
        return(NA)
      }
    }),
    stringsAsFactors = FALSE
  )
  
  nodes$group[is.na(nodes$group) | nodes$group == ""] <- "Unknown"
  nodes$id <- 0:(nrow(nodes) - 1)
  
  edges <- networkData %>%
    left_join(nodes, by = c("src" = "name")) %>%
    select(-src, -author) %>%
    rename(source = id) %>%
    left_join(nodes, by = c("target" = "name")) %>%
    select(-target, -author) %>%
    rename(target = id) %>%
    mutate(width = 1)
  
  nodes$oa_id <- nodes$name
  nodes$name <- nodes$author
  
  unique_groups <- unique(nodes$group)
  color_palette <- brewer.pal(min(length(unique_groups), 12), "Set3")
  
  ColourScale <- sprintf(
    'd3.scaleOrdinal().domain(["%s"]).range(["%s"]);',
    paste(unique_groups, collapse = '", "'),
    paste(color_palette, collapse = '", "')
  )
  
  openDOI <- "window.open(d.doi)"
  
  nwg <- forceNetwork(
    Links = edges,
    Nodes = nodes,
    Source = "source",
    Target = "target",
    NodeID = "name",
    Group = "group",
    Value = "width",
    opacity = 1,
    zoom = TRUE,
    colourScale = JS(ColourScale),
    fontSize = 20,
    legend = FALSE,
    clickAction = openDOI
  )
  
  nwg$x$nodes$doi <- nodes$doi
  
  # Create the legend HTML
  legend_html <- paste0(
    "<div style='margin: 20px 0;'>",
    "<h3>Legend</h3>",
    "<div style='display: flex; flex-wrap: wrap;'>"
  )
  
  for (i in seq_along(unique_groups)) {
    legend_html <- paste0(
      legend_html,
      "<div style='display: flex; align-items: center; margin-right: 15px; margin-bottom: 10px;'>",
      "<div style='width: 20px; height: 20px; background:", color_palette[i], "; margin-right: 8px;'></div>",
      "<span>", unique_groups[i], "</span>",
      "</div>"
    )
  }
  
  legend_html <- paste0(legend_html, "</div></div>")
  
  # Save the graph with the legend
  saveNetwork(nwg, file = file, selfcontained = TRUE)
  
  # Insert the legend into the saved HTML file
  graph_html <- readLines(file)
  insert_position <- grep("</body>", graph_html)[1] - 1
  graph_html <- append(graph_html, legend_html, after = insert_position)
  
  writeLines(graph_html, file)
  return(graph_html)
 # message("Graph with legend saved as: ", file)
}
 

# fn <- file.path(".", "interactive_graph_with_legend.html")
# if (file.exists(fn)) {
#     key_works <- readRDS(fn)
# } else {
# # Example call
plot_snowball_interactive(snowball, key_works)

```

## Chapter 2

```{r, eval=FALSE}

fn <- file.path(".", "data", "2_Chapter","chapter2_seed_works.rds")
if (file.exists(fn)) {
    key_works <- readRDS(fn)
} else {
    seeds <- fromJSON(file.path(".", "input", "2_Chapter","chapter2_seed.json"))
    #seeds$DOI[is.na(seeds$DOI)]  <- seeds$URL[is.na(seeds$DOI)] 
    key_works <- oa_fetch(
        entity = "works",
        doi=seeds$DOI[!is.na(seeds$DOI)],
        verbose = TRUE
    )
    saveRDS(key_works, fn)
}

ids <- openalexR:::shorten_oaid(key_works$id)

fn <- file.path(".", "data", "2_Chapter", "chapter2_snowball.rds")
if (file.exists(fn)) {
    snowball <- readRDS(fn)
} else {
    snowball <- oa_snowball(
        identifier = ids,
        verbose = TRUE
    )
    saveRDS(snowball, fn)
}

```

Number of key references: 162

Size of corpus: \~ 25,000

## Chapter 3

```{r, eval=FALSE}

fn <- file.path(".", "data", "3_Chapter","chapter3_seed_works.rds")
if (file.exists(fn)) {
    key_works <- readRDS(fn)
} else {
    seeds_1 <- fromJSON(file.path(".", "input", "3_Chapter","chapter3_1_seed.json"))
    seeds_2 <- fromJSON(file.path(".", "input", "3_Chapter","chapter3_2_seed.json"))
    seeds <- append(seeds_1,seeds_2)
    #seeds$DOI[is.na(seeds$DOI)]  <- seeds$URL[is.na(seeds$DOI)] 
    key_works <- oa_fetch(
        entity = "works",
        doi=seeds$DOI[!is.na(seeds$DOI)],
        verbose = TRUE
    )
    saveRDS(key_works, fn)
}

ids <- openalexR:::shorten_oaid(key_works$id)

fn <- file.path(".", "data", "3_Chapter", "chapter3_snowball.rds")
if (file.exists(fn)) {
    snowball <- readRDS(fn)
} else {
    snowball <- oa_snowball(
        identifier = ids,
        verbose = TRUE
    )
    saveRDS(snowball, fn)
}

```

Number of key references: 309

Size of corpus: \~ 30,000-35,000

## Chapter 4

```{r,eval=FALSE}

fn <- file.path(".", "data", "4_Chapter","chapter4_seed_works.rds")
if (file.exists(fn)) {
    key_works <- readRDS(fn)
} else {
    seeds <- fromJSON(file.path(".", "input", "4_Chapter","chapter4_seed.json"))
    #seeds$DOI[is.na(seeds$DOI)]  <- seeds$URL[is.na(seeds$DOI)] 
    key_works <- oa_fetch(
        entity = "works",
        doi=seeds$DOI[!is.na(seeds$DOI)],
        verbose = TRUE
    )
    saveRDS(key_works, fn)
}

ids <- openalexR:::shorten_oaid(key_works$id)

fn <- file.path(".", "data", "4_Chapter", "chapter4_snowball.rds")
if (file.exists(fn)) {
    snowball <- readRDS(fn)
} else {
    snowball <- oa_snowball(
        identifier = ids,
        verbose = TRUE
    )
    saveRDS(snowball, fn)
}

```

Number of key references: 85

Size of corpus: \~11,000

## Combined corpus from all chapters of MTA
